---
title: "timeFA Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Factor and Autoregressive Models for Tensor Time Series

The R package **timeFA** includes methods in our recent papers, including [Factor Models](https://arxiv.org/abs/1905.07530) and [Autoregressive Models](https://arxiv.org/abs/1812.08916) for High-Dimensional tensor Time Series. We demonstrate some examples for how to use our package here. To have more details please see the manual file for the full documentation.

#### Installation

Temporarily, we use GitHub to host this package source code then users can install the package directly from GitHub using the `install_github` function in the [devtools](https://github.com/r-lib/devtools#readme) package:
```{r, eval=FALSE}
library(devtools)
install_github("zebang/timeFA")
```


#### Autoregressive Models

We collect all estimation methods in one function `TAR`


#### Factor Models
